---
title: "hw_4_building_functions"
format: html
editor: visual
---

#### Define the contract

*Inputs*:

-   *ncount -* total number of individuals of each species (these inputs could be multiple values)

-   N -the total number of individuals of all species (this input is a single value calculated within the function)

*Output*: D - Simpson's Index, ranges between 0 and 1.

-   Lower values indicate lower diversity

*Parameters*: *ncount,* each number indicating the number of individuals observed for one species. Multiple values represent multiple species i.e., simpsons_index(c(15,20,30)) represents 3 species with 15, 20, and 30 individuals each.

#### Implement the function in R

```{r}
#' Simpson's Diversity Index 
#'
#' This function measures the probability that two individuals randomly selected from a sample will belong to the same species. 
#' @param ncount the number of individuals of a particular species (quantity)
#' @return Simpsons Index (D)

# function definition in-line
simpsons_index = function(ncount) {
  N = sum(ncount)
D = sum((ncount/N)^2)
return(D)
}

```

```{r}
simpsons_index(c(3,9))
```

```{simpsons_index(3,9)}


```
