---
title: "hw_4_building_functions"
format: html
editor: visual
---

#### Defining the contract

*Inputs*:

-   *ncount -* total number of individuals of each species (these inputs could be multiple values)

-   N -the total number of individuals of all species (this input is a single value calculated within the function)

*Output*: D - Simpson's Diversity Indices, ranges between 0 and 1.

-   Higher values indicate lower diversity (with this version of the equation)
-   0 represents infinite diversity
-   1 indicates no diversity

*Parameters*: *ncount,* each number indicating the number of individuals observed for one species. Multiple values represent multiple species.

-   i.e., `simpsons_index(c(15,20,30))` represents 3 species with 15, 20, and 30 individuals each

#### Implement the function in R

```{r}
#' Simpson's Diversity Index 
#'
#' This function measures the probability that two individuals randomly selected from a sample will belong to the same species. 
#' @param ncount the number of individuals of a particular species (quantity)
#' @return Simpsons Index (D)

# function definition in-line
simpsons_index = function(ncount) {
  N = sum(ncount)
D = sum((ncount/N)^2)
return(D)
}
```

#### Testing the Function

```{r}

# Test 1: Single species
simpsons_index(c(20))
## expected output D = 1 (no diversity)

# Test 2: Two species, equally distributed 
simpsons_index(c(20,20))
## expected output is a moderate D (0.5)

# Test 3: Randomly sampled data with 5 species 
set.seed(100)
species_counts <- sample(1:100, size = 5, replace = TRUE)
simpsons_index(species_counts)
## expected output - varies between sample data 
```
